Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A Linear Parameter Varying and Synergetic Control Approach

### 2017-10-29 paper/main_RA-L_v1_nadia.tex
Latest change of the day, finished with Section 5.

\begin{figure}[t]
\centering
\includegraphics[scale=.23,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}
\includegraphics[scale=.23,trim={0 0 0 0cm},clip]{With_Obstcle.png}
\caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl). The bottom example avoids a known obstacle in its workspace, while the top one does not.}
\label{fig:robot_example}
\end{figure}

@@

\newpage
\section{Experimental Validation} \label{Sec:Exp}
\subsection{JT-DS Learning Performance  Evaluation}
To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 7 different tasks (see Table \ref{tab:learning} and Fig. \ref{fig:behaviors}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space:
\begin{itemize}[leftmargin=*]
\item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
\item \textit{(3-5) Pouring motions:} We guide the robot to pour chips in a bowl with three different joint-space behaviors. The first is without an obstacle in the workspace. The second is with an obstacle in the upper hemisphere of the workspace, where the human is seen folding the elbow and lowering the shoulder of the robot arm to avoid the obstacle. The third instance is with an obstacle in the lower hemisphere, and now the human raises the shoulder to avoid it.
\item \textit{(6) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
\item \textit{(7) Singularity motions:} For this behavior we simulate joint trajectories in which the base rotates and all of the other joints are locked in singular configurations.
\end{itemize}

@@

@@

\begin{figure*}[!t]
	\vspace{-0.05cm}
	\begin{subfigure}[t]{0.42\linewidth}
		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
		\label{fig:snapshot:a}
	\end{subfigure}
	\begin{subfigure}[t]{0.567\linewidth}
		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
		\label{fig:snapshot:b}
	\end{subfigure}\\	
	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}.}
	\label{fig:snapshot}
	\vspace{-10pt}
\end{figure*}

@@

\subsubsection{Following Desired Joint and Task-Based Behaviors}
To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS  algorithm using  for behaviors (5) and (6).

The results can be seen in Fig. \ref{fig:snapshot:a} and Fig. \ref{fig:snapshot:b}. The JT-DS algorithm mimicked the joint-space behaviors of the demonstration (e.g. folding the elbow, lowering the shoulder), and therefore successfully avoided the obstacle while still converging to the desired Cartesian position. Meanwhile, SEDS only learned the demonstrated behavior in task-space meaning that the robot was not constrained in joint space. This ultimately led to one of its joints colliding with the obstacle. It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
. JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig. \ref{fig:foot}). This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.

%Further inspection reveals another advantage of JT-DS over stable Cartesian motion generators: not needing to preprocess demonstrations. In Cartesian-space dynamical systems, in order to guarantee stability, every demonstration must have the same target position. As a result, all demonstrated trajectories are warped to end at a single position, introducing a non-negligible difference between the original demonstrations and those used to train the DS. JT-DS, on the other hand, does not require any demonstration warping to learn. Rather than shifting demonstrations, the algorithm learns behavior regions with respect to the base frame of the manipulator. This effect can be seen in the diagrams in Fig. \ref{fig:Sin}, which show a comparison between JT-DS and a Cartesian-space DS trained through SEDS \cite{khansari2011learning}. The Cartesian-space system warps the demonstrations and as a result fails to track them, while JT-DS does not modify the demonstrations and thus succeeds.

\subsubsection{Systematic Assessment}
To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system. The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$. The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver. To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e. uniform behaviors.  The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse. Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero. Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.

\begin{table}[t]
	\centering
	\resizebox{\linewidth}{!}{\begin{tabular}{l|c|c|c|c}
		\hline \hline
		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
		Pour Obstacle 2 (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ 
		Pour Obstacle 2 (SEDS) &      7                &     3             &        -  &   -   \\		
		\hline
		Foot step  (JT-DS) &       3               &     1             &          1 & 0.01344   \\ 
		Foot step (SEDS)&       3               &     1             &          - & -   \\ \hline \hline
	\end{tabular}}
	\caption{Parameters and Performance of Comparative Experiments \textcolor{red}{(REDO)} }
	\label{table:1}
	\vspace{-10pt}
\end{table}

@@

\begin{table}[h]
	\vspace{0.1cm}
	\label{table:2}
	\centering
	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
		\hline\hline
		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
	\end{tabular}}
		\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$. The convergence duration is the time required to move within $0.001$m of the target. The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
\end{table}

\subsubsection{Transiting in singular configurations}
One of the main advantages of the proposed dynamical system is its ability to generate accurate paths in classical singular configurations. The first scenario, shown in Fig. \ref{fig:Sin}, was designed to assess this capability. The $10$ training demonstrations were constructed as follows: movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human demonstrator back-driving it. This restricted the motion to a series of arcs of different radii along the robot's motion boundary. In other words, the demonstrated motions were entirely within a classic kinematic singularity. Fig. \ref{fig:Sin} shows the \textit{demonstrated} motions and the motion  \textit{generated} by JT-DS \eqref{eq:ds}. The algorithm never requires the pseudo-inverse of the Jacobian matrix, so the generated motion perfectly follows the demonstrations throughout the workspace boundary.

\section{Discussion and Future Work} 
\label{Sec:Dis}
In this paper, we have presented a dynamical system in joint space that is provably Lyapunov-stable in task space and which replicates demonstrated joint-space behaviors. The desired motions are fast to compute, and smoothly handle singularities by avoiding the pseudo-inverse Jacobian. We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.

@@

%JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$. Previous work proposed adding discontinuous limits to JT control \cite{sciavicco1988solution}, and we additionally propose a continuous modification as follows: \textcolor{red}{Remove this part? Should we not talk about continuous joint limiting?} Replace the combined synergy matrix $A(q)$ with a joint-limited joint-behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that  
%\begin{equation}
%\label{eq:last_criteria}
%\begin{cases}
%\begin{split}
%& s^{i}(q^i_{max}) = 0&&&\\
%& s^{i}(q^i_{min})= 0&&&
%\end{split}
%\forall i\in 1,\dots,m
%\end{cases}
%\end{equation}
%$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t. the attractor. The structure of $S$ enforces kinematic joint limits because as a joint nears one of its limits, its magnitude decreases such that it will never actually collide with the limit. One way of constructing $S$ is to use the following function:
%\begin{equation}
%s^{i}(q) =
%1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{4}  ~\forall i\in 1,\dots,m\\
%\end{equation}
%The algorithm can similarly be expanded to accommodate joint velocity limits by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits. Specifically, given some set of velocity constraints for each joint $i$ s.t. $\dot{q}_i < \dot{q}_{i, max}$, we can design a new joint augmentation matrix $\mathcal{A}' = r \mathcal{A}$ where $r = \min \left ( 1, \max_i \left ( \frac{\dot{q}_{i, max}}{\dot{q}_i} \right ) \right )$. This velocity scalar would slow down the joint trajectory when the robot dynamics could not execute them, but would not change the position profile of the joint motion, preserving a major aspect of the joint behavior.

A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists. As mentioned in Sec. \ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^TP(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits. If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory. There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession. The smaller subtrajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits. Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.

@@

%Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system. This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors. \textcolor{red}{Are we actually still doing this?}

>>>>

\begin{figure}[t]
\centering
\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{With_Obstcle.png}
\caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl). The left example avoids a known obstacle in its workspace, while the right one does not.}
\label{fig:robot_example}
\vspace{-15pt}
\end{figure}

@@

\newpage
\section{Experimental Validation} \label{Sec:Exp}
\subsection{JT-DS Learning Performance  Evaluation}
To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 7 different tasks (see Table \ref{tab:learning} and Fig. \ref{fig:behaviors}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space:
\begin{itemize}[leftmargin=*]
\item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
\item \textit{(3-5) Pouring motions:} We guide the robot to pour chips in a bowl with three different joint-space behaviors. The first is without an obstacle in the workspace. The second is with an obstacle in the upper hemisphere of the workspace, where the human is seen folding the elbow and lowering the shoulder of the robot arm to avoid the obstacle. The third instance is with an obstacle in the lower hemisphere, and now the human raises the shoulder to avoid it.
\item \textit{(6) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
\item \textit{(7) Singularity motions:} Movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human. This restricted the motion to a series of arcs of different radii along the robot's motion boundary. 
\end{itemize}

@@

\begin{figure*}[ht]
	\vspace{-0.05cm}
	\begin{subfigure}[t]{0.42\linewidth}
		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
		\label{fig:snapshot:a}
	\end{subfigure}
	\begin{subfigure}[t]{0.567\linewidth}
		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
		\label{fig:snapshot:b}
	\end{subfigure}\\	
	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}.}
	\label{fig:snapshot}
	\vspace{-10pt}
\end{figure*}

@@

@@

\subsubsection{Following Desired Joint and Task-Based Behaviors}
We compare tracking capabilities of our JT-DS method with those of SEDS+IK for behaviors \textit{(5)} and \textit{(6)}. In Fig. \ref{fig:snapshot:a} and \ref{fig:snapshot:b} we can qualitatively see the difference between these two approaches for behavior \textit{(5)}. The JT-DS algorithm mimicked the joint-space behavior of the demonstration (e.g. folding the elbow, raising the shoulder), successfully avoiding the obstacle while still converging to the desired Cartesian position. Meanwhile, SEDS+IK only learned the demonstrated behavior in task-space, it does not have the capabilities of constraining motion in joint-space. This ultimately led to one of its joints colliding with the obstacle. It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position. For the foot-step behavior \textit{(6)} JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig. \ref{fig:snapshot:c} and \ref{fig:foot}). This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.

@@

\subsubsection{Systematic Assessment}
To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a SEDS+IK, we simulated $400$ simple motions on each system. The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$. The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver. To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e. uniform behaviors.  The results, summarized in Table \ref{tab:sys}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse. As the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero. Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than SEDS+IK.

\begin{table}[t]
	\label{table:2}
	\centering
	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
		\hline\hline
		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
	\end{tabular}}
		\caption{Each simulated trajectory is initialized at $q=[0~\dots~0]^T$. The convergence duration is the time required to move within $0.001$m of the target. The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.\label{tab:sys}}
		\vspace{-10pt}
\end{table}

\subsubsection{Transiting through singular configurations}
One of the main advantages of the proposed DS is its ability to generate accurate paths in classical singular configurations. To evaluate this we generated behavior \textit{(7)}. The $10$ training demonstrations were constructed as follows:

In other words, the demonstrated motions were entirely within a classic kinematic singularity. Fig. \ref{fig:Sin} shows the \textit{demonstrated} motions and the motion  \textit{generated} by JT-DS \eqref{eq:ds}. The algorithm never requires the pseudo-inverse of the Jacobian matrix, so the generated motion perfectly follows the demonstrations throughout the workspace boundary.

\section{Discussion} 
\label{Sec:Dis}
In this paper, we have presented a dynamical system in joint space that is provably Lyapunov-stable in task space and which replicates demonstrated joint-space behaviors. The desired motions are fast to compute, and smoothly handle singularities by avoiding the pseudo-inverse Jacobian. We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.

@@

A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists. As mentioned in Sec. \ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^TP(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits. If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory. There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession. The smaller sub-trajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits. Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.

@@
